---
title: "Homework 3 Joseph Shin"
author: "Joseph Shin"
date: "2025-09-04"
output:
  bookdown::pdf_document2:
    latex_engine: lualatex
  tufte::tufte_handout:
    latex_engine: xelatex
---

EXPLANATIONS ARE IN THE OTHER PDF FILE


```{r}
library(outliers)
library(ggplot2)
crime = read.csv("uscrime.txt", sep = '\t')
crime_outlier = crime[,16]
plot(density(crime_outlier))
hist(crime_outlier, freq = FALSE)
shapiro.test(crime_outlier)
```

Looking at this histogram of the number of crimes, the distribution is not normal, but instead, is right skewed. Additionally, according to the Shapiro test, the p-value is less than 0.05, so we can reject the null hypothesis and conclude that the distribution is not normal. Additionally, in the density curve, you can see the right skew. The right skew may mean that there are potential outliers in the higher values. The grubbs.test() is performed under the assumption that the distribution is normal. Since it is not normal, we may have to include different methods to identify the outliers. However, we can still use a for loop to run a grubbs.test to identify one outlier, remove it, and then run the grubbs.test() until the distribution becomes normal.


Use the grubbs function to take out one outlier, and then test the grubbs function again, and keep doing it until there are no outliers left.

```{r}
removed_outliers <- c()

repeat {
  # Test for normality
  p_normality <- shapiro.test(crime_outlier)$p.value
  
  if (p_normality > 0.05) {
    cat("Data is now approximately normal. Stopping.\n")
    break
  }
  
  # Run Grubbs' test
  g_test <- grubbs.test(crime_outlier, type = 10)
  print(g_test)
  
  # Extract the outlier
  outlier_val <- as.numeric(strsplit(g_test$alternative, " ")[[1]][3])
  
  if (!is.na(outlier_val)) {
    cat("Removing outlier:", outlier_val, "\n")
    removed_outliers <- c(removed_outliers, outlier_val)
    crime_outlier <- crime_outlier[crime_outlier != outlier_val]
  } else {
    cat("No more outliers found. Stopping.\n")
    break
  }
}
```

```{r}
bp_crime = boxplot(crime_outlier)
```

According to the boxplot, the maximum value is 1,635. However, there are three values higher than the maximum that are potential outliers: 1674, 1969, and 1993. We can use the boxplot outlier formula to see if these points are considered outliers or not. Since Q3 = 1057.5 and Q1 = 658, the IQR = Q3 - Q1 = 399.5. Then, we use the equation Q3 + 1.5 * IQR = 1057.5 + 1.5 * 399.5 = 1656.75. Since 1675, 1969, and 1993 are all greater than 1635, we can consider these three points outliers. Based on the boxplot, there does not seem to be an outlier that is lower than the minimum. Therefore, I decided to use type 20 because we would test two outliers in one tail (higher side).



```{r}
temps = read.delim('temps.txt', check.names = FALSE)
temps
```
Using July through October daily-high-temperature data for Atlanta for 1996 through 2015, use a CUSUM approach to identify when unofficial summer ends (i.e., when the weather starts cooling off) each year

```{r, warning = FALSE}
cusum = function(temps, target = mean(temps), k = 0.3) {
  n = length(temps)
  pos = numeric(n)
  neg = numeric(n)
  
  for (i in 2:n) {
    pos[i] = max(0, pos[i-1] + temps[i] - target - k)
    neg[i] = max(0, neg[i-1] + target - temps[i] - k)
  }
  list(pos = pos, neg = neg)
}

years = colnames(temps)[-1]

cusum_results = list()

for (year in years) {
  year_data = temps[[year]]
  result = cusum(year_data)
  cusum_results[[year]] = result
}

#Calculating the average of the temperature where the peak is at so that I can use that as my threshold.

peak_vals <- sapply(cusum_results, function(x) max(x$pos))
mean_peak <- mean(peak_vals)
h = quantile(peak_vals, probs = 0.15)
h


for (year in years) {
  c_pos <- cusum_results[[year]]$pos
  c_neg <- cusum_results[[year]]$neg
  
  plot(c_pos, type = "l", col = "red", lwd = 2,
       main = paste("Year:", year),
       ylab = "CUSUM", xlab = "Day",
       ylim = c(min(c_neg), max(c_pos)))  # allow full range
  
  lines(c_neg, col = "blue", lwd = 2)
  abline(h = 0, lty = 2)  # baseline
  
  legend("topright", legend = c("CUSUM+", "CUSUM−"),
         col = c("red", "blue"), lwd = 2, cex = 0.3)
  abline(h = 262, col = "red", lty = 2)
  abline(v = which.max(cusum_results[[year]]$pos), col = "blue", lty = 2)
}
```

I set my k-value to 0.3. The reason for this is that I did not want to include meaningless deviations that may cause C+ to increase. At the same time, since a couple graphs were barely above the threshold, I wanted to keep k at a balanced value so that it showed the clear result of when "unofficial summer" ended. I only wanted to consider the temperature of the days where it was meaningfully above the average. For my threshold, I took the average of the peak values of each year. However, the threshold was still pretty strict, so I took the 15% quantile of that average, and ended up using 262 as my threshold. The x-axis, Days, is a time series that goes through each day of the data. The y-axis, C+, represents the accumulated amount of above-average temperatures over time. Since I perform the calculation in my for loop (temp - mean - k), it created the deviation for me, and then I would add up those deviations to represent the C+ accumulation. Once the decrease reached the threshold, that would be the day I would consider the end of "unofficial summer". Almost every plot dropped below the threshold from Days 80-100. This means that the end of unofficial summer is roughly between September 18th to October 8th. Additionally, the solid blue line shows the increase of the cooling temperatures. In almost every plot, the blue line increases when the red line hits the threshold. This proves that "unofficial summer" is ending. However, there were a couple plots whose max accumulated temperature didn't reach my threshold. However, I still looked at the plots and noticed the same pattern in the drop of accumulated temperatures and an increase in the C- blue line from September 18th to October 8th.




```{r}
years = colnames(temps)[-1]
summer_average = sapply(temps[ , years], mean, na.rm = TRUE)
names(summer_average) = years
```

```{r}
cusum_clim = function(temps, target = mean(temps), k = 0, reset = FALSE) {
  n = length(temps)
  cumsum = numeric(n)
  for (i in 2:n) {
    cumsum[i] = cumsum[i-1] + (temps[i] - target - k)
  if (reset && cumsum[i] < 0) cumsum[i] = 0
  }
  return (cumsum)
}
```

```{r, warning = FALSE}
cusum_val = cusum_clim(summer_average)

plot(1996:2015, cusum_val, type = 'l', col = 'darkred', lwd = 2, xlab = 'Year', ylab = 'CUSUM of Summer Avg Temp.', main = 'Detecting Warming in Atlanta Summers')
abline(h = 0, lty = 2)
```
I set k = 0 because I want take all deviations into account. Since we are focusing on the years, small shifts in those annual averages are also meaningful. Therefore, I want to capture all changes. Looking at the graph, the slope suddenly increases starting from 2009. Therefore, the change point is at year 2009. From there on, it strictly increases and peaks at year 2012. The dashed line at h = 0 means that there was no cumulative deviation from the mean. When the line is below the dashed line, that means that there is still a cooling trend since the CUSUM is in the negatives. When it is above the dashed line, that means that there is a warming trend. The CUSUM is above h = 0 around year 2012. Therefore, there is a clear warming trend that year.


```{r}
n = length(summer_average)
diff_means = numeric(n - 1)

for (i in 1:(n - 1)) {
  before = summer_average[1:i]
  after = summer_average[(i + 1):n]
  diff_means[i] = abs(mean(after) - mean(before))
}
```

```{r}
cpt_index = which.max(diff_means)
cpt_year = years[cpt_index]
```

```{r, warning = FALSE}
plot(years, summer_average, type = "b", pch = 19, col = "darkorange",
     xlab = "Year", ylab = "Avg Summer Temp",
     main = "Average Summer Temps in Atlanta (1996–2015)")

abline(v = cpt_year, col = "blue", lty = 2)

```

The graph with the orange line shows the difference in average temperature for each year. The blue line clearly shows that the change point starts at year 2009. There is a huge average difference between 2009 and 2010. It is about a 6 degree difference. After that, the average difference is on a decline.

