---
title: "Homework 2"
author: "Joseph Shin"
date: "2025-08-25"
output:
  pdf_document: default
  html_document: default
---

Question 3.1: Using the same data set (credit_card_data.txt or credit_card_data-headers.txt) as in Question 2.2, use the ksvm or kknn function to find a good classifier:
(a) using cross-validation (do this for the k-nearest-neighbors model; SVM is optional); and
(b) splitting the data into training, validation, and test data sets (pick either KNN or SVM; the other
is optional).


```{r}
library(kernlab)
library(kknn)
```

```{r}
data = read.csv("credit_card_data-headers copy.txt", header = TRUE, sep = "\t")
```

Question 3.1 a) Cross Validation

```{r}
set.seed(1)

random = sample(1:nrow(data), as.integer(0.7*nrow(data)))

training_data = data[random,]
testing_data = data[-random,]

train.kknn(R1 ~., data = training_data, kmax = 50, scale = TRUE)

training_prediction = rep(0, (nrow(training_data)))

training_accuracy = 0

for (i in 1:nrow(training_data)) {
  model = kknn(R1 ~., training_data[-i,], training_data[i,], k = 41, kernel = "optimal", scale = TRUE)
  training_prediction[i] = as.integer(fitted(model) + 0.5)
}

training_accuracy = sum(training_prediction == training_data[,11]) / nrow(training_data)

testing_prediction = rep(0,(nrow(testing_data)))
testing_accuracy = 0

for (i in 1:nrow(testing_data)) {
  model = kknn(R1 ~., testing_data[-i,], testing_data[i,], k = 41, kernel = "optimal", scale = TRUE)
  testing_prediction[i] = as.integer(fitted(model) + 0.5)
}

testing_accuracy = sum(testing_prediction == testing_data[,11]) / nrow(testing_data)

training_accuracy

testing_accuracy
```

Question 3.1 b) Splitting into training, validation, and testing data sets

Training data: 60%, Validation data: 20%, Testing data: 20%

```{r}
set.seed(1)

training_index = sample(1:nrow(data), as.integer(0.6 * nrow(data)))
remaining_index = setdiff(1:nrow(data), training_index)
validation_index = sample(remaining_index, as.integer(0.5 * length(remaining_index)))
testing_index = setdiff(remaining_index, validation_index)

training_data = data[training_index, ]
validation_data = data[validation_index, ]
testing_data = data[testing_index, ]
```

```{r}
training_prediction = rep(0, (nrow(training_data)))
training_accuracy = 0
i = 0
accuracytab = data.frame(matrix(nrow = 25, ncol = 2))
colnames(accuracytab) = c("K", "Accuracy")
for (X in 1:25) {
  for (i in 1:nrow(training_data)) {
    model = kknn(R1 ~. , training_data[-i, ], training_data[i, ], k = X, kernel = "optimal", scale = TRUE)
    training_prediction[i] = as.integer(fitted(model) + 0.5)
  }
  
  training_accuracy = sum(training_prediction == training_data[,11]) / nrow(training_data)
  accuracytab[X, 1] = X
  accuracytab[X, 2] = training_accuracy
}
accuracytab

plot(accuracytab[,1], accuracytab[,2])
```
```{r}
validation_prediction = rep(0, (nrow(validation_data)))
validation_accuracy = 0
i = 0
accuracytab_valid = data.frame(matrix(nrow = 25, ncol = 2))
colnames(accuracytab_valid) = c("K", "Accuracy")
for (X in c(5:8, 21)) {
  for (i in 1:nrow(validation_data)) {
    model = kknn(R1 ~. , validation_data[-i, ], validation_data[i, ], k = X, kernel = "optimal", scale = TRUE)
    validation_prediction[i] = as.integer(fitted(model) + 0.5)
  }
  
  validation_accuracy = sum(validation_prediction == validation_data[,11]) / nrow(validation_data)
  accuracytab_valid[X, 1] = X
  accuracytab_valid[X, 2] = validation_accuracy
}
accuracytab_valid

plot(accuracytab_valid[,1], accuracytab_valid[,2])
```



```{r}
testing_prediction = rep(0, (nrow(testing_data)))
testing_accuracy = 0
i = 0
accuracytab_testing = data.frame(matrix(nrow = 25, ncol = 2))
colnames(accuracytab_testing) = c("K", "Accuracy")
for (X in 1:25) {
  for (i in 1:nrow(testing_data)) {
    model = kknn(R1 ~. , testing_data[-i, ], testing_data[i, ], k = X, kernel = "optimal", scale = TRUE)
    testing_prediction[i] = as.integer(fitted(model) + 0.5)
  }
  
  testing_accuracy = sum(testing_prediction == testing_data[,11]) / nrow(testing_data)
  accuracytab_testing[X, 1] = X
  accuracytab_testing[X, 2] = testing_accuracy
}
accuracytab_testing

plot(accuracytab_testing[,1], accuracytab_testing[,2])
```


Question 4.2: Clustering

```{r}
library(ggplot2)
library(tibble)
library(factoextra)
library(gridExtra)
iris = read.table('iris.txt', header = TRUE)
iris_clean = scale(iris[,-5])
```



```{r}
km2 = kmeans(iris_clean, centers = 2, nstart = 20)
km3 = kmeans(iris_clean, centers = 3, nstart = 20)
km4 = kmeans(iris_clean, centers = 4, nstart = 20)
km5 = kmeans(iris_clean, centers = 5, nstart = 20)

plot2 = fviz_cluster(km2, geom = "point", data = iris_clean) + ggtitle("Two Clusters")
plot3 = fviz_cluster(km3, geom = "point", data = iris_clean) + ggtitle("Three Clusters")
plot4 = fviz_cluster(km4, geom = "point", data = iris_clean) + ggtitle("Four Clusters")
plot5 = fviz_cluster(km5, geom = "point", data = iris_clean) + ggtitle("Five Clusters")

grid.arrange(plot2, plot3, plot4, plot5, nrow = 2, top = "Figure 1: Cluster Plots")
```

```{r}
table(km2$cluster, iris$Species)
table(km3$cluster, iris$Species)
table(km4$cluster, iris$Species)
table(km5$cluster, iris$Species)
```


```{r}
fviz_nbclust(iris_clean, kmeans, method = "wss") +
  geom_vline(xintercept = 3, linetype = 2) +
  labs(subtitle = "Figure 2")
```

```{r}
accuracy_kmeans = list()
accuracyAll = (50+39+36) / 150
accuracy_kmeans[1] = accuracyAll
cat("Prediction Accuracy with all Predictors: ", as.integer(accuracyAll * 100), "% \n")
```

```{r}
predictor_function = function(x1, x2, x3) {
  predictor_kmeans = kmeans(iris_clean[,c(x1, x2, x3)], 3, nstart = 20)
  table(predictor_kmeans$cluster, iris$Species)
}
```

```{r}
predictor_function(1, 2, 3)

predictor_function(1, 2, 4)

predictor_function(2, 3, 4)

predictor_function(1, 3, 4)

```

```{r}
accuracy1 = (49 + 35 + 37) / 150

accuracy2 = (49 + 38 + 34) / 150

accuracy3 = (49 + 42 + 38) / 150

accuracy4 = (50 + 44 + 36) / 150

c(accuracy1, accuracy2, accuracy3, accuracy4)
```

The best combination is the fourth accuracy percentage, which values at 86.66667%. This would mean that the best combination of predictors is sepal length, petal length, and petal width.




