---
title: "ISYE 6501 HW 1"
author: "Joseph Shin"
date: "2025-08-19"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(kernlab)
```

```{r}
data = read.csv("credit_card_data-headers.txt", header = TRUE, sep = "\t")
```

2.2.1: Using the vanilladot kernel

C = 100

```{r}
model = ksvm(as.matrix(data[,1:10]), as.factor(data[,11]), type = "C-svc", kernel = "vanilladot", C = 100, scaled = TRUE)
model
a = colSums(model@xmatrix[[1]] * model@coef[[1]])
a
a0 = -model@b
a0
pred = predict(model, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

y = factor(data[,11])
false_positives <- sum(pred == 1 & y == 0)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm = table(actual = y, pred = pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```

C = 1000000

```{r}
model2 = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "vanilladot", C = 1000000, scaled = TRUE)
model2
a = colSums(model2@xmatrix[[1]] * model2@coef[[1]])
a0 = -model2@b
a0
pred = predict(model2, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm <- table(actual=y, pred=pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```


C = .01

```{r}
model3 = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "vanilladot", C =.01, scaled = TRUE)
model3
a = colSums(model3@xmatrix[[1]] * model3@coef[[1]])
a
a0 = -model3@b
a0
pred = predict(model3, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm <- table(actual=y, pred=pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```


C = .001

```{r}
model4 = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "vanilladot", C = .001, scaled = TRUE)
model4
a = colSums(model4@xmatrix[[1]] * model4@coef[[1]])
a
a0 = -model4@b
a0
pred = predict(model4, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm <- table(actual=y, pred=pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```

C = .0001

```{r}
model5 = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "vanilladot", C = .0001, scaled = TRUE)
model5
a = colSums(model5@xmatrix[[1]] * model5@coef[[1]])
a
a0 = -model5@b
a0
pred = predict(model5, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)
```

2.2.2

Testing different kernels (I will be testing different C values to try to increase the accuracy percentage of each model as well as minimizing the number of false positives.)

Using rbfdot kernel:

```{r}
rbfmod = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "rbfdot", C = 1, scaled = TRUE)
rbfmod
a = colSums(rbfmod@xmatrix[[1]] * rbfmod@coef[[1]])
a
a0 = -rbfmod@b
a0
pred = predict(rbfmod, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm <- table(actual=y, pred=pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```

Using polydot kernel:

```{r}
polymod = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "polydot", C = .001, scaled = TRUE)
polymod
a = colSums(polymod@xmatrix[[1]] * polymod@coef[[1]])
a
a0 = -polymod@b
a0
pred = predict(polymod, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm <- table(actual=y, pred=pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```


Using tanhdot kernel:

```{r}
tanhmod = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "tanhdot", C = 100, scaled = TRUE)
tanhmod
a = colSums(tanhmod@xmatrix[[1]] * tanhmod@coef[[1]])
a
a0 = -tanhmod@b
a0
pred = predict(tanhmod, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm <- table(actual=y, pred=pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```


Using anovadot kernel:

```{r}
anovamod = ksvm(as.matrix(data[,1:10]), data[,11], type = "C-svc", kernel = "anovadot", C = .01, scaled = TRUE)
anovamod
a = colSums(anovamod@xmatrix[[1]] * anovamod@coef[[1]])
a
a0 = -anovamod@b
a0
pred = predict(anovamod, data[,1:10])
pred
sum(pred == data[,11]) / nrow(data)

x = as.matrix(data[,1:10])
y = factor(data[,11], levels = c(0,1))
cm <- table(actual=y, pred=pred)
FP <- cm["0","1"]  # false positives: predicted 1, actual 0
FN <- cm["1","0"]  # false negatives: predicted 0, actual 1
FP; FN
```


Question 2.2.3: k-nearest-neighbors classification

```{r}
library(kknn)
```

```{r}
accuracy_knn = function(x) {
  pred = rep(0, (nrow(data)))
  
for (i in 1:nrow(data)) {
  
model = kknn(R1 ~. , train = data[-i,], test = data[i,], k = x, scale = TRUE, kernel = "optimal")

pred[i] <- as.integer(fitted(model)+0.5)

}
  
  accuracy = sum(pred == data[,11]) / nrow(data)
  return(accuracy)
}

accuracy_vector <- 1:20
for (x in 1:20){
  accuracy_vector[x] = accuracy_knn(x) * 100
}
accuracy_vector

knn_value = c(1:20)
plot(knn_value, accuracy_vector)
```

```{r}
max(accuracy_vector)
which.max(accuracy_vector)
```


Rectangular Kernel

```{r}
accuracy_knn = function(x) {
  pred = rep(0, (nrow(data))) #Create a vector of 0's to store in the predicted
  
for (i in 1:nrow(data)) {
  
model = kknn(R1 ~. , train = data[-i,], test = data[i,], k = x, scale = TRUE, kernel = "rectangular")

pred[i] <- as.integer(fitted(model)+0.5)

}
  
  accuracy = sum(pred == data[,11]) / nrow(data)
  return(accuracy)
}

accuracy_vector <- 1:20
for (x in 1:20){
  accuracy_vector[x] = accuracy_knn(x) * 100
}
accuracy_vector

knn_value = 1:20
plot(knn_value, accuracy_vector)
```

```{r}
max(accuracy_vector)
which.max(accuracy_vector)
```






