---
title: "Homework 5"
author: "Joseph Shin"
date: "2025-09-19"
output:
  pdf_document: default
  html_document: default
---

```{r}
crime_data = read.table('uscrime.txt', header = TRUE, sep = '\t')
crime_data
```
Use regression (a useful R function is lm or glm) to predict the observed crime rate in a city with the following data:

M = 14.0
So = 0
Ed = 10.0
Po1 = 12.0
Po2 = 15.5
LF = 0.640
M.F = 94.0
Pop = 150
NW = 1.1
U1 = 0.120
U2 = 3.6
Wealth = 3200
Ineq = 20.1
Prob = 0.04
Time = 39.0
Show your model (factors used and their coefficients), the software output, and the quality of fit.
Note that because there are only 47 data points and 15 predictors, you’ll probably notice some
overfitting. We’ll see ways of dealing with this sort of problem later in the course.

```{r}
crime_data = crime_data[-c(4, 26),]
crime_data
```
# From Homework 3, I removed two outliers from this exact dataset. The outliers that I removed were when the crime rate was at 1993 and 1969. Therefore, I chose to remove those two observations in this dataset before I perform regression because it may interfere with my analysis.


# For AIC, we should use the corrected AIC because it is a small dataset. Look for smaller AIC for better fit.




```{r}
library(stats)
lin_mod = lm(Crime ~ ., data = crime_data)
summary(lin_mod)
confint(lin_mod, level = 0.95)
```
# If a confidence interval does not include 0, then that means that the predictor IS significant.

Looking at the summary of the linear model of all the predictors, the coefficients that are significant (for alpha = 0.05) are Ed, Ineq, and Prob. The adjusted R-squared is 51.14%, and the model's p-value is 0.001618. Looking at the confidence intervals of the predictors (95% confidence), the significant predictors are Ed, Ineq, and Prob, which are the same compared to the summary. If the confidence interval did not include 0, then that would mean that the predictor is not significant. Therefore, I was able to identify which predictors were significant based on this rule.


```{r}
lin_aic = AIC(lin_mod)
lin_bic = BIC(lin_mod)
lin_aic
lin_bic
```
The linear model's BIC is 654.5932.

# Corrected AIC formula: AICc = AIC + 2k(k+1) / (n-k-1) where k = number of parameters being estimated and n = number of data points

```{r}
n = length(lin_mod$fitted.values)
k = length(coef(lin_mod))
lin_aicc = lin_aic + (2 * k * (k + 1)) / (n - k - 1)
lin_aicc
```
The linear model's Corrected AIC is 596.725.

```{r}
library(car)
vif(lin_mod)
```


```{r}
library(car)
vif(lin_mod)
```

Let's try removing the VIF's that cause multicollinearity. Therefore, I will try removing Po2 since it seems like it's correlated with Po1 and Wealth, which seems like it's correlated with Ineq.

```{r}
test_lin_mod = lm(Crime~. -Wealth - Po2, data = crime_data)
summary(test_lin_mod)
vif(test_lin_mod)
```
```{r}
lin_aic1 = AIC(test_lin_mod)
lin_bic1 = BIC(test_lin_mod)
lin_bic1

#Corrected AIC

n = length(test_lin_mod$fitted.values)
k = length(coef(test_lin_mod))
lin_aicc1 = lin_aic1 + (2 * k * (k + 1)) / (n - k - 1)
lin_aicc1
```

After performing VIF on the new model, all of the variables VIF's are close to 5 and lower, which is a good sign. That shows that there is no correlation between the predictor variables, so there is also no interruption with the individual effects to the response variable, Crime.

Based on the p-value tests of each predictor, the significant predictors are Ed, Po1, U2, Ineq, and Prob. Therefore, I will try creating a model with those variables.

```{r}
test_lin_mod1 = lm(Crime ~ Ed + Po1 + U2 + Ineq + Prob, data = crime_data)
summary(test_lin_mod1)
```

```{r}
lin_aic2 = AIC(test_lin_mod1)
lin_bic2 = BIC(test_lin_mod1)
lin_bic2

#Corrected AIC

n = length(test_lin_mod1$fitted.values)
k = length(coef(test_lin_mod1))
lin_aicc2 = lin_aic2 + (2 * k * (k + 1)) / (n - k - 1)
lin_aicc2
```


As we can see here, the only insignificant predictor is U2. Let's try taking out that variable one last time.

```{r}
test_lin_mod2 = lm(Crime ~ Ed + Po1 + Ineq + Prob, data = crime_data)
summary(test_lin_mod2)
vif(test_lin_mod2)
confint(test_lin_mod2)
```

```{r}
lin_aic3 = AIC(test_lin_mod2)
lin_bic3 = BIC(test_lin_mod2)
lin_bic3

#Corrected AIC

n = length(test_lin_mod2$fitted.values)
k = length(coef(test_lin_mod2))
lin_aicc3 = lin_aic3 + (2 * k * (k + 1)) / (n - k - 1)
lin_aicc3
```

Here is all the information about this model. The P-value is 3.554e-07, which is the lowest that I have received out of all of the linear models I have created. The Adjusted R-squared is 53.88%, which means that 53.88% of the variation is explained by the model with the variables: Ed, Po1, Ineq, and Prob. This current model's Adjusted R-Squared is only slightly less than the other models (maximum 2-3%). The VIF of the predictors are all below 5, which is very good because that means that multicollinearity is not present. All of the predictors are significant because all of their p-values are lower than 0.05. Finally, the Corrected AIC is lower than that of the previous model (619.8326 < 621.8009). The BIC is lower than that of the previous model, but only by ~3 (629.1341 < 632.237). Therefore, this current BIC is 'somewhat likely' to do better than the previous model's. Based on all this information, I believe that this is the best performing model for lm().

Now, we will predict the observed crime rate using the information in the summary of this model.



```{r}
test_point = data.frame(M = 14.0, So = 0, Ed = 10.0, Po1 = 12.0, Po2 = 15.5, LF = 0.640, M.F = 94.0, Pop = 150, NW = 1.1, U1 = 0.120, U2 = 3.6, Wealth = 3200, Ineq = 20.1, Prob = 0.04, Time = 39.0)

pred_lin_model = predict(test_lin_mod2, test_point)
round(pred_lin_model)
```

The predicted observed crime rate from the model that consists of Ed, Po1, Ineq, and Prob, is 1239. This means that there were 1239 number of offenses per 100,000 population in 1960.

```{r}
library(DAAG)
set.seed(42)
lm_cv = cv.lm(crime_data, test_lin_mod2, n = 4)
```

The MSE is 70591.57, which is close to the lowest MSE out of all of the linear models that I have created.










Now, let's test out different models using glm().


# Poisson model

```{r}
poisson_mod = glm(Crime~., data = crime_data, family = poisson)
summary(poisson_mod)
confint(poisson_mod)
```

```{r}
poisson_aic = AIC(poisson_mod)
poisson_bic = BIC(poisson_mod)
poisson_aic
poisson_bic
```
# Poisson Model Corrected AIC

```{r}
n = length(poisson_mod$fitted.values)
k = length(coef(poisson_mod))
poisson_aicc = poisson_aic + (2 * k * (k + 1)) / (n - k - 1)
poisson_aicc
```
#Pseudo R-Squared for Poisson Model

```{r}
library(pscl)
pR2(poisson_mod)
```
#Poisson cv.glm()
```{r}
library(boot)
poisson_cv = cv.glm(crime_data, poisson_mod, K = 4)
poisson_cv

mse_raw = poisson_cv$delta[1]
mse_adjusted = poisson_cv$delta[2]
c(mse_raw, mse_adjusted)
```


```{r}
vif(poisson_mod)
```

```{r}
poisson_mod1 = glm(Crime ~. - Po2 - Wealth - NW - Time, data = crime_data, family = poisson)
summary(poisson_mod1)
```

```{r}
poisson_cv1 = cv.glm(crime_data, poisson_mod1, K = 4)
poisson_cv1

mse_raw = poisson_cv1$delta[1]
mse_adjusted = poisson_cv1$delta[2]
c(mse_raw, mse_adjusted)
```

```{r}
poisson_aic1 = AIC(poisson_mod1)
poisson_bic1 = BIC(poisson_mod1)
poisson_aic1
poisson_bic1

n = length(poisson_mod1$fitted.values)
k = length(coef(poisson_mod1))
poisson_aicc1 = poisson_aic1 + (2 * k * (k + 1)) / (n - k - 1)
poisson_aicc1
```


```{r}
pR2(poisson_mod1)
```



Comparing the two poisson models that I have created, the original poisson model has better traits. The first model has a lower BIC, lower corrected AIC, a slightly larger R-Squared (from the McFadden), and a lower MSE. The only thing is the first model includes all of the predictors, and some predictors are insignificant while the other model's predictors are all significant. If we're looking for predictive accuracy, we would use the first model. Therefore, I will predict the values according to the first model.

```{r}
test_point = data.frame(M = 14.0, So = 0, Ed = 10.0, Po1 = 12.0, Po2 = 15.5, LF = 0.640, M.F = 94.0, Pop = 150, NW = 1.1, U1 = 0.120, U2 = 3.6, Wealth = 3200, Ineq = 20.1, Prob = 0.04, Time = 39.0)

pred_poisson_model = predict(poisson_mod, test_point, type = 'response')
pred_poisson_model
```

This prediction value looks highly unreasonable, considering the fact that the poisson model is performed on a different scale. Therefore, the multiple linear regression model is a better model to use.















